---
title: "PML"
output: html_document
---
Load required libraries and set environment variables
```{r, echo=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(rattle)
library(randomForest)
```
Read files that are already downloaded to local drive
```{r}
trainRaw <- read.csv("pml-training.csv")
testRaw <- read.csv("pml-testing.csv")

dim(trainRaw)
dim(testRaw)

```
## Cleaning data

Step 1 : Removing Missing Values

```{r}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 

```
Step 2 : Removing date that will not be used 
```{r}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

## Slice the data
Cleaned training sliced in to 
  Training data set (70%) and
  Validation data set (30%).
Validation data set is to conduct cross validation at a later stage.

```{r}
set.seed(22519) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling

Random Forest algorithm is used to fit a predictive model for activity recognition   because it automatically selects important variables and is robust to correlated covariates & outliers in general. A 5-fold cross validation when applying the algorithm.

```{r}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf

predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)

accuracy <- postResample(predictRf, testData$classe)
accuracy

oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose

```
So, the estimated accuracy of the model is 99.35% and the estimated out-of-sample error is 0.65%.

## Using Using ML algorithms for prediction of Test Data Set (Randon Forest)
```{r}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result

```

## Using Using ML algorithms for prediction of Test Data Set (RandomTree)
```{r}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) 
```

