---
title: "Practical Machine Learning - Project"
output: html_document
---
By Jana

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

The goal of this project is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants (Adelmo, Carlitos, Charles, Eurico, Jeremy and Pedro) and predict the manner in which they did the exercise. Thes particpants were asked to perform barbell lifts correctly and incorrectly in 5 different ways (Stitting, Stitting Down, Standing, Standing Up and Walking) which in given as the "classe" variable in the training set. 

## Data Sourne
The training data was downloaded from 
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
  
The test data was downloaded from 
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
  
Load required libraries and set environment variables
```{r, echo=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(rattle)
library(randomForest)
```
Read files that are already downloaded to local drive
```{r}
trainRaw <- read.csv("pml-training.csv")
testRaw <- read.csv("pml-testing.csv")

dim(trainRaw)
dim(testRaw)

```
## Cleaning data

Step 1 : Removing Missing Values

```{r}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 

```
Step 2 : Removing date that will not be used 
```{r}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

## Slice the data
Cleaned training sliced in to 
  Training data set (70%) and
  Validation data set (30%).
Validation data set is to conduct cross validation at a later stage.

```{r}
set.seed(22519) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling

Random Forest algorithm is used to fit a predictive model for activity recognition   because it automatically selects important variables and is robust to correlated covariates & outliers in general. A 5-fold cross validation when applying the algorithm.

```{r}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf

predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)

accuracy <- postResample(predictRf, testData$classe)
accuracy

oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose

```
So, the estimated accuracy of the model is 99.35% and the estimated out-of-sample error is 0.65%.

## Using Using ML algorithms for prediction of Test Data Set and creating files
```{r}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
answers <- result

pml_write_files <- function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i], file=filename, quote=FALSE,row.names=FALSE, col.names=FALSE)
  }
}
pml_write_files(answers)



```

## Using Using ML algorithms for prediction of Test Data Set (RandomTree)
```{r}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) 
fancyRpartPlot(treeModel)



```

